---
title: "Raw OX-I Performance and Data Quality"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---

What effect do raw ratio means and variability have on data quality?

Test value and variance of raw 12C current, 14/12 ratio and 13/12 ratio vs
normalized SD of standards. First section looks at within-target variance. 
Second section looks at the variability of standards within a wheel.

## Get and process data

```{r}
suppressMessages(library(amstools))
library(odbc)
suppressMessages(library(tidyverse))
theme_set(theme_bw())

# a plotting function
cplot <- function(data, x, y) {
ggplot(data, aes_string(x, y, color = "system")) + 
    geom_point() + 
    geom_smooth(method="lm") + 
    facet_grid(.~ system, scales = "free_x")
}
```


Get current, raw ratios, and normalized ratios for standards within a date range

```{r}
date <- "2019-05-01"

sql <- paste0("SELECT wheel, runtime, tp_num, sample_name, 
	         le12c, he13_12, he14_12
               FROM snics_raw
               WHERE sample_type = 'S'
		    AND runtime > '", date, "'
		    AND ok_calc = 1")

sqlres <- paste0("SELECT wheel, target.tp_num, target.osg_num, runtime, gf_date, sample_name, 
		    fm_corr, sig_fm_corr
                  FROM snics_results
                  JOIN target ON target.tp_num = snics_results.tp_num
                  JOIN graphite ON graphite.osg_num = target.osg_num
                  WHERE sample_type = 'S'
		    AND runtime > '", date, "'")
          
con <- conNOSAMS()
raw <- dbGetQuery(con, sql)
res <- dbGetQuery(con, sqlres)
```

Summarize raw data by target.

```{r}
rawsum <- raw %>% 
	mutate(he14_12 = he14_12*1E12, 
	       le12c = le12c*1E6) %>% 
       group_by(tp_num) %>% 
       summarise(le12c.m = mean(le12c),
		 le12c.s = sd(le12c),
		 he13_12.m = mean(he13_12),
		 he13_12.s = sd(he13_12),
		 he14_12.m = mean(he14_12),
		 he14_12.s = sd(he14_12))

data <- inner_join(res, rawsum, by = "tp_num") %>% 
	mutate(system = toupper(substring(wheel, 1, 5)), 
	       le12c.m = ifelse(system == "USAMS", -le12c.m, le12c.m),
	       rt = as.POSIXct(runtime),
	       age = gf_date - rt)

# remove flyers and oxII
data <- filter(data, fm_corr > 0.9, fm_corr < 1.2, 
	       sig_fm_corr < 0.005, he14_12.s < 0.03, le12c.m < 150,
	       osg_num > 151250,
	       he13_12.m > 0.9)
```


## Within target variability 

Examine means and variability of raw ratios and normalized standards by
target and separate by system.
```{r}
data %>% group_by(system) %>% 
	summarize(Fm = mean(fm_corr),
		  Fm.sd = sd(fm_corr),
		  he1412 = mean(he14_12.m),
		  he1412.sd = sd(he14_12.m),
		  he1312 = mean(he13_12.m),
		  he1312.sd = sd(he13_12.m),
		  le12c = mean(le12c.m),
		  le12c.sd = sd(le12c.m)) %>%
	knitr::kable(digits = 4)

```

### Plots of raw ratios and within-target variance by OSG number.

```{r}
ggplot(data, aes(gf_date, he13_12.m, color = system)) +
  geom_point() + facet_grid(system ~ ., scales = "free_y")
ggplot(data, aes(gf_date, he13_12.s, color = system)) +
  geom_point() + facet_grid(system ~ ., scales = "free_y")
ggplot(data, aes(gf_date, he14_12.m, color = system)) +
  geom_point() + facet_grid(system ~ ., scales = "free_y")
ggplot(data, aes(gf_date, he14_12.s, color = system)) +
  geom_point() + facet_grid(system ~ ., scales = "free_y")
ggplot(data, aes(gf_date, fm_corr, color = system)) +
  geom_point() + facet_grid(system ~ ., scales = "free_y")
```

```{r}
ggplot(data, aes(age, he13_12.m, color = system)) +
  geom_point() + geom_smooth(method="lm") +
  facet_grid(system ~ ., scales = "free_y")
ggplot(data, aes(age, he14_12.m, color = system)) +
  geom_point() + geom_smooth(method="lm") +
  facet_grid(system ~ ., scales = "free_y")
ggplot(data, aes(age, fm_corr, color = system)) +
  geom_point() + geom_smooth(method="lm") +
  facet_grid(system ~ ., scales = "free_y")
```

### Raw ratio mean and sd vs normalized fm

Raw data and within target variabilty vs normalized fm and
standard deviation of fm. First, any effect of raw ratios on 
precision of individual targets? Looks like nothing too obvious.

Note that fits of a "ball" of data are pretty unreliable.

```{r}
ggplot(data, aes(le12c.m, he14_12.m, color = system)) +
  geom_point() + geom_smooth(method="lm") +
  facet_grid(system ~ ., scales = "free_y")
ggplot(data, aes(le12c.m, he14_12.s, color = system)) +
  geom_point() + geom_smooth(method="lm") +
  facet_grid(system ~ ., scales = "free_x")
ggplot(data, aes(le12c.m, fm_corr, color = system)) +
  geom_point() + geom_smooth(method="lm") +
  facet_grid(system ~ ., scales = "free_y")
ggplot(data, aes(le12c.m, sig_fm_corr, color = system)) +
  geom_point() + geom_smooth(method="lm") +
  facet_grid(system ~ ., scales = "free_x")
cplot(data, "he13_12.m", "he14_12.m")
ggplot(data, aes(he13_12.m, he14_12.m, color = system)) +
  geom_point() + geom_smooth(method="lm") 
cplot(data, "he13_12.m", "he14_12.s")
cplot(data, "he13_12.m", "fm_corr")
cplot(data, "he13_12.m", "sig_fm_corr")
cplot(data, "he13_12.s", "fm_corr")
cplot(data, "he13_12.s", "sig_fm_corr")
cplot(data, "he14_12.m", "he14_12.s")
cplot(data, "he14_12.m", "fm_corr")
cplot(data, "he14_12.m", "sig_fm_corr")
cplot(data, "he14_12.s", "fm_corr")
cplot(data, "he14_12.s", "sig_fm_corr")
```

Slope and intercept of 14/12 vs 13/12 for CFAMS and USAMS

```{r}

data %>% 
    group_by(system) %>% 
    do({
      mod = lm(he14_12.m ~ he13_12.m, data = .)
      data.frame(Intercept = coef(mod)[1],
                 Slope = coef(mod)[2])
    })
```

## Summarize and plot by wheel 

Summarizing is neccessary to get SD's by wheel, but I think this tends to blur
relationships due to variability in graphite.

```{r}
wheelsum <- data %>% group_by(wheel, system) %>% 
	summarise(date = mean(as.Date(runtime)),
	          fm_corr.m = mean(fm_corr),
		  fm_corr.sd = sd(fm_corr),
		  le12c.me = mean(le12c.m),
		  le12c.s = sd(le12c.m),
		  he13_12.me = mean(he13_12.m),
		  he13_12.s = sd(he13_12.m),
		  he14_12.me = mean(he14_12.m),
		  he14_12.s = sd(he14_12.m))

wheelsum <- filter(wheelsum, fm_corr.sd < 0.01)
wheelsum %>% group_by(system) %>% 
	summarize(Fm = mean(fm_corr.m),
		  Fm.sd = mean(fm_corr.sd),
		  he1412 = mean(he14_12.me),
		  he1412.sd = mean(he14_12.s),
		  he1312 = mean(he13_12.me),
		  he1312.sd = mean(he13_12.s),
		  le12c = mean(le12c.me),
		  le12c.sd = sd(le12c.me)) %>%
	knitr::kable(digits = 4)
```

### Plots by date

```{r}
ggplot(wheelsum, aes(date, he13_12.me, color = system)) +
  geom_point() + facet_grid(system ~ ., scales = "free_y")
ggplot(wheelsum, aes(date, he13_12.s, color = system)) +
  geom_point() + facet_grid(system ~ ., scales = "free_y")
ggplot(wheelsum, aes(date, he14_12.me, color = system)) +
  geom_point() + facet_grid(system ~ ., scales = "free_y")
ggplot(wheelsum, aes(date, he14_12.s, color = system)) +
  geom_point() + facet_grid(system ~ ., scales = "free_y")
ggplot(wheelsum, aes(date, fm_corr.m, color = system)) +
  geom_point() + facet_grid(system ~ ., scales = "free_y")
ggplot(wheelsum, aes(date, fm_corr.sd, color = system)) +
  geom_point() + facet_grid(system ~ ., scales = "free_y")
```

### Plots of raw data vs corrected fm by wheel

plots of raw data effect on sd of standards by wheel

```{r}

# a plotting function
cplotns <- function(data, x, y) {
ggplot(data, aes_string(x, y, color = "system")) + 
    geom_point() + 
    facet_grid(.~ system, scales = "free_x")
}
cplotns(wheelsum, "le12c.me", "fm_corr.sd")
cplotns(wheelsum, "he14_12.me", "fm_corr.sd")
cplotns(wheelsum, "he13_12.me", "fm_corr.sd")
cplotns(wheelsum, "he14_12.s", "fm_corr.sd")
cplotns(wheelsum, "he13_12.s", "fm_corr.sd")
```
